Stack Safety as a Security Property

### 2021-02-17 main.tex
Redid a bunch of trace stuff. Needs to be clearer on whether observations are in the state, but intro is next on my plate

\judgmenttwo{\(\mach_1 \stepstoobs{\obs} \mach_2\)}{\((\mach_1, \pol_1)
               \polstep \pol_2\)}
            {\(\mpstate{\mach_1}{\pol_1} \stepstopol{\obs}
               \mpstate{\mach_2}{\pol_2}\)}

@@

\judgmenttwo{\(\mpstate{\mach_1}{\pol_1} \stepstopol{\obs} \mpstate{\mach_2}{\pol_2}\)}
            {\(\mach_1, \context_1 \constep \context_2\)}
            {\(\mpcstate{\mach_1}{\pol_1}{\context_1} \stepstocon{\obs}
               \mpcstate{\mach_2}{\pol_2}{\context_2}\)}

@@

The ``trace-of'' operator, written \(\mach \hookrightarrow \machT\),
coinductively relates a machine state with the trace of machine states
and observations produced by repeated application of step:

\judgmenttwo{\(\mach_0 \stepstoobs{\obs} \mach_1\)}{\(\mach_1 \hookrightarrow \machT\)}
            {\(\mach_0 \hookrightarrow (\mach_0,\obs) \machT\)}%
%
\noindent
(We use juxtaposition of an element and a sequence to represent ``cons.'')
Since the step function is total, if \(\mach \hookrightarrow \machT\),
then \(\machT\) must be infinite.

@@

In a similar fashion, we define \(\mpstatename \hookrightarrow \MPT\)
to relate a machine-policy state pair to the MP-trace
induced by the step function on $\MPS$. Unlike the $\MOTRACE$ above, this trace can
be finite as the policy can cause a fault, in which case the final observation is
a \(\tau\):

\begin{minipage}{.4\textwidth}
\judgment{\(\neg \exists \pol_1 . \mpstate{\mach_0}{\pol_0} \polstep \pol_1\)}
         {\(\mpstate{\mach_0}{\pol_0} \hookrightarrow \mpostate{\mach_0}{\pol_0}{\tau}\)}
\end{minipage}
\begin{minipage}{.55\textwidth}
  \judgmenttwo{\(\mpstate{\mach_0}{\pol_o} \stepstopol{\obs}
                 \mpstate{\mach_1}{\pol_1}\)}
              {\(\mpstate{\mach_1}{\pol_1} \hookrightarrow \MPT\)}
              {\(\mpstate{\mach_0}{\pol_0} \hookrightarrow
                 \mpostate{\mach_0}{\pol_0}{\obs} \MPT\)}
\end{minipage}

We will also use \(\pi_m\) to project the trace of machine states
out of a trace, \(\pi_p\) to project the trace of policy states, and
\(\pi_o\) to project the observations, with \(\pi_m\) and \(\pi_o\) overloaded
for machine traces and MP-traces.

@@

\paragraph*{Until}
The operation \(\PUT~(f,\Trace)\)
%\bcp{Wouldn't
%``While'' be more mnemonic?} \sna{Appealing, but it's more of a
%``while-not''? Trying out ``until''. In any case... moving away
%from doing joins means this needs to get some work to eliminate
%the trailing observation. Here's an attempt.}
takes a trace
%% \bcp{I changed a : here to $\in$.  We should check if there are
%% any more to fix.}
\(\Trace \in \TRACE{\TRACEELEMS} \times \OBSS\) and a predicate
on elements \(f \subset \TRACEELEMS\) and gives the prefix of
\(\Trace\) ending with the first element on which \(f\) holds.
One subtlety is that, because the observation associated with some element
in a trace
represents what can be observed when this element is executed (i.e., the
observation on that step), when we take a prefix from a trace containing
observations
we must replace the final observation with a \(\tau\).

  \begin{center}
  \begin{minipage}{.3\textwidth}
    \judgment{\(f\ \traceelem\)}
             {\(\PUT~(f,(\traceelem,\obs)\Trace) = (\traceelem,\tau)\)}
  \end{minipage}
%
  \begin{minipage}{.3\textwidth}
  \judgment{}
           {\(\PUT~(f,(\traceelem,\obs)) = (\traceelem,\tau)\)}
%
  \end{minipage}
  \begin{minipage}{.3\textwidth}
  \judgmenttwo{\(\neg f\ \traceelem\)}{\(\PUT~(f,\Trace) = \Trace'\)}
              {\(\PUT~(f,(\traceelem,\obs)\Trace) = (\traceelem,\obs)\Trace'\)}
  \end{minipage}
  \end{center}
%
If the resulting trace is a strict prefix of the parameter, then the
predicate must hold on its final element. If the predicate never holds, then
\(\PUT\) is the identity function.

@@

We will frequently take a prefix of the trace from an initial state up
through the first state where some condition holds on the machine state.
This can easily be implemented using \(\PUT\), and we provide special notation
for convenience.
%
If \(f\) a predicate on machine states, we define \(\mach
\hookrightarrow \machT | f\) (read ``\(\machT\) is the prefix of
running \(\mach\) up to \(f\)'') and similarly for traces with policies:

\begin{center}
\begin{minipage}{.3\textwidth}
\judgmenttwo{\(\mach \hookrightarrow \machT'\)}
            {\(\PUT~(f,\machT') = \machT\)}
            {\(\mach \hookrightarrow \machT | f\)}
\end{minipage}
\begin{minipage}{.6\textwidth}
\judgmenttwo{\(\mpstatename \hookrightarrow \MPT'\)}
            {\(\PUT~(\lambda \mpstate{\mach}{\pol}.f\ \mach,\MPT') = \MPT\)}
         {\(\mpstatename \hookrightarrow \MPT | f\)}
\end{minipage}
\end{center}

%    \COQ{this operation steps until it reaches a state on which \(f\) holds,
%    and is infinite if it doesn't. It doesn't compute the observation of the step after the final,
%    which coq would. So coq may need a ``step until'' relation rather than ``prefix up to''.
%    Still thinking about this.}\leo{So this is strictly weaker than the one we have in Coq, right?
%      So its only the Eager implies Lazy direction that might need tweaking, but not the Test implies Eager one?}
%    \leo{Also, it steps until it reaches an f-state or if the policy raises a fault}

@@

\[
(\push{\domMap}{n}{\components}) ~ \component =
\begin{cases}
  \sealed{\depth} & \text{if } \domMap ~ \component = \unsealed
                    \text{ and } \component \in \components \\
  \domMap ~ \component & \text{else} \\
\end{cases}
\]
\[
(\pop{\domMap}{n}) ~ \component =
\begin{cases}
  \unsealed & \text{if } \domMap ~ \component = \sealed{\depth-1} \\
  \domMap ~ \component & \text{else} \\
\end{cases}
\]

@@

\begin{center}
\begin{minipage}{.4\textwidth}
  \judgmenttwo{\(\mach \stepsto \mach'\)}
              {\(\codemap ~ (\mach ~ \PCname) = \funid, \mathbf{Some} ~ \callmap\)}
              {\(\mach,(\domMap,\depth) \stepsto_C \mach',(\push{\domMap}{\depth}{\sealcon ~ \mach},\depth+1)\)}
\end{minipage}
\hspace*{0.1\textwidth}
\begin{minipage}{.4\textwidth}
  \judgmenttwo{\(\mach \stepsto \mach'\)}
              {\(\codemap ~ (\mach ~ \PCname) = \funid, \mathbf{Some} ~ \retmap\)}
              {\(\mach,(\domMap,\depth) \stepsto_C \mach',(\pop{\domMap}{\depth},\depth-1)\)}
\end{minipage}
\end{center}

\begin{center}
  \judgment[Default]
           {\(\mach \stepsto \mach'\)}
           {\(\mach,\context \stepsto_C \mach',\context\)}
\end{center}

@@

\definition A system enjoys {\em ultra-eager integrity} if, for any initial state \(\mach_0\),
its initial context \(\context_0\), and its initial policy state \(\pol_0\), and each
pair of adjacent states \(\mach,(\domMap,\depth),\pol \stepsto_C \mach',\context',\pol'\) in
the induced trace from \(\mach_0,\context_0,\pol_0\), if \(\domMap ~ \component = \sealed{\depth'}\)
for any \(\depth'\), then \(\mach(\component) = \mach'(\component)\).

@@

\definition Call Rule goes here.

\definition A system enjoys {\em eager integrity} if, for any initial state \(\mach_0\), initial
context \(\context_0\), and initial policy state \(\pol_0\) and any call depth \(\depth\), each segment
\(\MCPTrace\) taken from the induced trace from \(\mach_0,\context_0,\pol_0\) by
\(\prop (\domMap,\depth') = \depth' \geq \depth\) satisfies \(\pi_m(\head ~ \MCPTrace) \component =
\pi_m(\last ~ \MCPTrace) \component\) for all \(\component\) where \(\domMap ~ \component = \sealed{\depth''}\)
for any \(\depth''\).

@@

\definition A system enjoys {\em eager confidentiality} if, for any initial state \(\mach_0\), initial
context \(\context_0\), and initial policy state \(\pol_0\) and any call depth \(\depth\), each segment
\(\MCPTrace\) taken from the induced trace from \(\mach_0,\context_0,\pol_0\) by
\(\prop (\domMap,\depth') = \depth' \geq \depth\) behaves the same as any \(\components\)-variant
as follows.

\begin{itemized}
\item
\item
\item
\end{itemized}

>>>>

\judgmenttwo{\(\mach \stepstoobs{\obs} \mach'\)}{\((\mach, \pol)
               \polstep \pol'\)}
            {\(\mpstate{\mach}{\pol} \stepstopol{\obs}
               \mpstate{\mach'}{\pol'}\)}

@@

\judgmenttwo{\(\mpstate{\mach}{\pol} \stepstopol{\obs} \mpstate{\mach'}{\pol'}\)}
            {\(\mach, \context \constep \context'\)}
            {\(\mpcstate{\mach}{\pol}{\context} \stepstocon{\obs}
               \mpcstate{\mach'}{\pol'}{\context'}\)}

@@

The ``trace-of'' operator, written \(\mach \hookrightarrow \machT\),
coinductively relates a machine state with the trace of machine states
and observations produced by repeated application of step. It's defined in
terms of a coinductive helper, \(\mathit{obs-forward}\):

\judgmenttwo{\(\mach \stepstoobs{\obs'} \mach'\)}
            {\(\mathit{obs-forward} ~ \mach' ~ \obs' = \machT\)}
            {\(\mathit{obs-forward} ~ \mach ~ \obs = (\mach,\obs) \machT\)}%
%
\judgment{\(\mathit{obs-forward} ~ \mach ~ \tau = \machT\)}
         {\(\mach \hookrightarrow \machT\)}
\noindent
(We use juxtaposition of an element and a sequence to represent ``cons.'')
Since the step function is total, if \(\mach \hookrightarrow \machT\),
then \(\machT\) must be infinite.

@@

In a similar fashion, we define \(\mpcstatename \hookrightarrow \MPCT\)
to relate a machine-policy-context state triple to the MPCO-trace
induced by the step function.

\judgment{\(\neg \exists \pol' . \mpcstate{\mach}{\pol} \polstep \pol'\)}
         {\(\mathit{obs-forward} ~ \mpcstate{\mach}{\pol}{\context} ~ \obs =
           \mpcstate{\mach}{\pol}{\context},\obs\)}
%
\judgmenttwo{\(\mpcstate{\mach}{\pol}{\context} \stepstoobs{\obs'} \mpcstate{\mach'}{\pol'}{\context'}\)}
            {\(\mathit{obs-forward} ~ \mpcstate{\mach'}{\pol'}{\context'} ~ \obs' = \machT\)}
            {\(\mathit{obs-forward} ~ \mpcstate{\mach}{\pol}{\context} ~ \obs =
              (\mpcstate{\mach}{\pol}{\context},\obs) \machT\)}%
%
\judgment{\(\mathit{obs-forward} ~ \mpcstatename ~ \tau = \machT\)}
         {\(\mpcstatename \hookrightarrow \machT\)}

We will also use \(\pi_m\), \(\pi_p\), \(\pi_c\), and \(\pi_o\) to project the components out of an
MPC state, and overload them to project a trace of the components out of an MPC trace.

@@

\paragraph*{Until}
The operation \(\PUT ~ f ~ \Trace\) takes a trace
\(\Trace \in \TRACE{\TRACEELEMS \times \OBSS}\) and a predicate
on elements \(f \subset \TRACEELEMS\) and gives the prefix of
\(\Trace\) ending with the first element on which \(f\) holds.

\begin{center}
  \begin{minipage}{.3\textwidth}
    \judgment{\(f ~ \traceelem\)}
             {\(\PUT ~ f ~ \traceelem\Trace = \traceelem\)}
  \end{minipage}
%
  \begin{minipage}{.3\textwidth}
    \judgment{}
             {\(\PUT ~ f ~ \traceelem = \traceelem\)}
%
  \end{minipage}
  \begin{minipage}{.3\textwidth}
    \judgmenttwo{\(\neg f ~ \traceelem\)}{\(\PUT ~ f ~ \Trace = \Trace'\)}
                {\(\PUT ~ f ~ \traceelem\Trace = \traceelem\Trace'\)}
  \end{minipage}
\end{center}
%
If the resulting trace is a strict prefix of the parameter, then the
predicate must hold on its final element. If the predicate always holds, then
\(\PUT\) is the identity function.

@@

We will frequently take a prefix of the trace from an initial state up
through the first state where some condition holds on the machine state.
This can easily be implemented using \(\PUT\), and we provide special notation
for convenience.
%
If \(f\) is a predicate on machine states, we define \(\mpcstatename
\hookrightarrow \MPCT | f\) (read ``\(\MPCT\) is the prefix of
running \(\mpcstatename\) up to \(f\)'') and similarly for traces with policies:

\begin{center}
\judgmenttwo{\(\mpcstatename \hookrightarrow \MPCT'\)}
            {\(\PUT ~ f ~ \mpcstatename ~ \MPCT' = \MPCT\)}
            {\(\mpcstatename \hookrightarrow \MPCT | f\)}
\end{center}

\paragraph*{Context Segments}

In particular, we wish to formulate properties around consecutive subtraces
in which the context obeys a given predicate. This will commonly be used to extract
individual calls from a trace by taking subtraces in which the stack is
at or below a given depth. If \(f \subseteq \MACHS \times \CONTEXTS\), then
we have \(\mpcstatename \hookrightarrow \MPCT \langle f \rangle\), pronounced
``\(\mpcstatename\) segmented by \(f\) yields a subtrace \(\MPCT\).''

\begin{center}
  \judgmenttwo[Now]
              {\(f ~ \mach ~ \context\)}
              {\(\mach,\pol,\context \hookrightarrow \MPCT | \neg f\)}
              {\(\mach,\pol,\context \hookrightarrow \MPCT \langle f \rangle\)}
%
  \judgmentthree[Skip]
                {\(f ~ \mach ~ \context\)}
                {\(\mach,\pol,\context \hookrightarrow \MPCT | \neg f\)}
                {\(\last(\MPCT) \hookrightarrow \MPCT' \langle f \rangle\)}
                {\(\mach,\pol,\context \hookrightarrow \MPCT' \langle f \rangle\)}
%
  \judgmentthree[Search]
                {\(\neg f ~ \mach ~ \context\)}
                {\(\mach,\pol,\context \hookrightarrow \MPCT | f\)}
                {\(\last(\MPCT) \hookrightarrow \MPCT' \langle f \rangle\)}
                {\(\mach,\pol,\context \hookrightarrow \MPCT' \langle f \rangle\)}
\end{center}

\paragraph*{Reachability}

We often want to quantify over all traces from any initial state so that our
properties apply to the system as a whole. We say that a state \(\mpcstatename'\)
is {\em reachable} if for some initial state \(\mpcstatename\), \(\mpcstatename \hookrightarrow \MPCT\)
and \(\MPCT\) contains \(\mpcstatename'\). A {\it reachable segment} \(\MPCT \langle f \rangle\) is a subtrace
such that from some initial state \(\mpcstatename\), \(\mpcstatename \hookrightarrow \MPCT \langle f \rangle\).

@@

\[
(\push{\domMap}{\depth}{\components}) ~ \component =
\begin{cases}
  \sealed{\depth} & \text{if } \domMap ~ \component = \unsealed
                    \text{ and } \component \in \components \\
  \domMap ~ \component & \text{else} \\
\end{cases}
\]
\[
(\pop{\domMap}{\depth}) ~ \component =
\begin{cases}
  \unsealed & \text{if } \domMap ~ \component = \sealed{\depth-1} \\
  \domMap ~ \component & \text{else} \\
\end{cases}
\]

@@

\begin{center}
\begin{minipage}{.4\textwidth}
  \judgmenttwo{\(\mach \stepsto \mach'\)}
              {\(\codemap ~ (\mach ~ \PCname) = \funid, \mathbf{Some} ~ \callmap\)}
              {\(\mach,(\domMap,\depth) \stepstocon \mach',(\push{\domMap}{\depth}{\sealcon ~ \mach},\depth+1)\)}
\end{minipage}
\hspace*{0.1\textwidth}
\begin{minipage}{.4\textwidth}
  \judgmenttwo{\(\mach \stepsto \mach'\)}
              {\(\codemap ~ (\mach ~ \PCname) = \funid, \mathbf{Some} ~ \retmap\)}
              {\(\mach,(\domMap,\depth) \stepstocon \mach',(\pop{\domMap}{\depth},\depth-1)\)}
\end{minipage}
\end{center}

\begin{center}
  \judgment[Default]
           {\(\mach \stepsto \mach'\)}
           {\(\mach,\context \stepstocon \mach',\context\)}
\end{center}

@@

\definition A system enjoys {\em ultra-eager integrity} if for any reachable state
\(\mpcstatename\), \(\mpcstatename'\) such that \(\mpcstatename \stepstocon{} \mpcstatename'\),
and component \(\component\), if \(\pi_\context(\mpcstatename) = \domMap,\_\) and
\(\domMap ~ \component = \sealed{\_}\), then \(\pi_\mach(\mpcstatename) ~ \component =
\pi_\mach(\mpcstatename') ~ \component\).

@@

\definition Let \(\mpcstatename\) be a reachable state whose next instruction is annotated
as a call and whose call depth is \(\depth\). Let \(f ~ \mach ~ \context\) hold when \(\context = \_,\depth\),
and \(\components\) be the set of components \(\component\) such that
\(\sealcon ~ \pi_\mach(\mpcstatename) ~ \component\) holds. Then the {\em Call Rule} says that if
\(\mpcstatename \stepstocon \mpcstatename'\), it is sound to assume that
\(\mpcstatename' \hookrightarrow \MPCT | f\) and if \(\last(\MPCT) = \mach'\), then
\(\pi_\mach(\mpcstatename) ~ \component = \mach' ~ \component\).

\definition Let \(\depth\) be any depth and \(f_\depth ~ \mach ~ (\_, \depth')\) hold when
\(\depth' \geq \depth\). A system enjoys {\em eager integrity} if each reachable segment
\(\MPCT \langle f_\depth \rangle\) satisfies \(\pi_m(\head(\MPCT)) ~ \component =
\pi_m(\last(\MPCT)) ~ \component\) for all \(\component\) where \(\domMap ~ \component =
\sealed{\_}\).

@@

\definition Let \(\depth\) be any depth and \(f_\depth ~ \mach ~ (\_, \depth')\) hold for
\(\depth' \geq \depth\). For any reachable segment \(\MPCT \langle f_\depth \rangle\),
let \(\head(\MPCT) = \mach,\pol,(\domMap,\depth),\_\) and
\(\components\) be the set of components \(\component\) such that \(\domMap ~ \component \not = \outside\).
Then take any \(\nach\) such that \(\mach \approx_\components \nach\). \(\MPCT\) corresponds to the
trace of a subroutine from its call to its return.

If we take the induced trace from this variant until a return,
\(\nach,\pol,\context \hookrightarrow \NPCT | \neg f_\depth\), we get a subtrace corresponding
to the same subrouting called with the same (register) arguments but with a potentially different
state at the call-site. A system enjoys {\em eager confidentiality} if for all such pairs of traces,
they correspond according to one of four cases:

\begin{enumerate}
\item \(\MPCT\) may be finite and \(\pi_\context(\last(\MPCT)) = (\_,\depth')\) with \(\depth' < \depth\),
  and \(\pi_\context(\last(\NPCT)) = (\_,\depth')\) with \(\depth' < \depth\). That is, both traces end
  having left the call of depth \(\depth\). Further:
  \begin{itemize}
  \item They have the same observable behavior, \(\pi_\obs(\MPCT) \simeq \pi_\obs(\NPCT)\)
  \item Their final states have the same differences, \(\pi_\mach(\head(\MPCT)), \pi_\mach(\last(\MPCT)) \diamond
  \pi_\mach(\head(\MPCT)), \pi_\mach(\last(\NPCT))\)
  \end{itemize}
\item \(\MPCT\) and \(\NPCT\) may both be infinite, with \(\pi_\obs(\MPCT) \simeq \pi_\obs(\NPCT)\)
\item \(\MPCT\) may be finite but end without returning -- which can only have happened if it fail-stopped --
  and its observable behavior prefixes that of \(\NPCT\), \(\pi_\obs(\MPCT) \lesssim \pi_\obs(\NPCT)\)
\item \(\NPCT\) may end with a fail-stop as in case 3, with \(\pi_\obs(\NPCT) \lesssim \pi_\obs(\MPCT)\)
\end{enumerate}
